# Analyze Spatial: `sf`

## Overview {.objectives}

**Questions**

- How to elegantly conduct complex spatial analysis by chaining operations?
- What is the percent area of water by region across the United States?

**Objectives**

- Use the `%>%` operator (aka "then" or "pipe") to pass output from one function into input of the next.
- Calculate metrics on spatial attributes.
- Aggregate spatial data with metrics.
- Display a map of results.

## Prerequisites {.prereq}

**R Skill Level**: Intermediate - you've got basics of R down.

You will use the `sf` package for vector data along with the `dplyr` package for calculating and manipulating attribute data.

```{r tidy-load-packages, message=F, warning=F}
# load packages
library(tidyverse)  # load dplyr, tidyr, ggplot2 packages
library(sf)         # vector reading & analysis

# set working directory to data folder
# setwd("pathToDirHere")
```

## States: read and plot

Similar to [Lesson 9: Handling Spatial Projection & CRS in R](https://jsta.github.io/R-spatial-raster-vector-lesson/09-vector-when-data-dont-line-up-crs/), we'll start by reading in a polygon shapefile using the `sf` package. Then use the default `plot()` function to see what it looks like.

```{r plot-states}
library(here)

states_shp <- here("data/neon-us-boundary/US-State-Boundaries-Census-2014.shp")

# read in states
states <- read_sf(states_shp)

# plot the states
plot(states)
```

Notice the default plot on `sf` objects outputs colorized values of the first 9 of 10 columns. Use the suggestion from the warning to plot the 10th column.

```{r plot-states-10}
# plot 10th column
plot(states, max.plot = 10)

# show columns of the data frame
names(states)

# look at table
glimpse(states)

# convert to tibble for nicer printing
as_tibble(states)

names(states)

# inspect the class(es) of the states object
class(states)
```

The class of the `states` object is both a simple feature (`sf`) as well as a data frame, which means the many useful functions available to a data frame (or "tibble") can be applied.

To plot the column of interest, feed the "slice" of that column to the `plot()` function.

```{r plot-states-region}
plot(states['region'])
```

**Question**: To motivate the spatial analysis for the rest of this lesson, you will answer this question: "_**What is the percent water by region?**_"

## Challenge: analytical steps? {.challenge}

Outline a sequence of analytical steps needed to arrive at the answer.

### Answers {.solution} 

1. **Sum** the area of water (`AWATER`) and land (`ALAND`) per region.
1. **Divide** the area of water (`AWATER`) by the area of land  (`ALAND`) per region to arrive at percent water.
1. Show **table** of regions sorted by percent water.
1. Show **map** of regions by percent water with a color ramp and legend.

## Regions: calculate % water

- Use the `%>%` operator (aka "then" or "pipe") to pass output from one function into input of the next.
    - In RStudio, see menu Help > Keyboard Shortcuts Help for a shortcut to the "Insert Pipe Operator".
    
- Calculate metrics on spatial attributes.
    - In RStudio, see menu Help > Cheatsheets > [Data Manipulation with dplyr, tidyr](https://github.com/rstudio/cheatsheets/raw/master/source/pdfs/data-transformation-cheatsheet.pdf).

- Aggregate spatial data with metrics.

```{r calc-regions}
regions = states %>%
  group_by(region) %>%
  summarize(
    water = sum(AWATER),
    land  = sum(ALAND)) %>%
  mutate(
    pct_water = water / land * 100 %>% round(2))

# object
regions
```

Notice the geometry in the column. To remove the geometry column pipe to `st_set_geometry(NULL)`. To arrange in descending order use `arrange(desc(pct_water))`.

```{r regions-table}
# table
regions %>%
  st_set_geometry(NULL) %>%
  arrange(desc(pct_water))
```

## Regions: plot

Now plot the regions.

```{r plot-regions-pctwater}
# plot, default
plot(regions['pct_water'])
```

## Regions: ggplot

The `ggplot2` library can [visualise sf objects](http://ggplot2.tidyverse.org/reference/ggsf.html).

- In RStudio, see menu Help > Cheatsheets > [Data Visualization with ggplot2](https://github.com/rstudio/cheatsheets/raw/master/source/pdfs/ggplot2-cheatsheet-2.1.pdf).

```{r ggplot-regions-pctwater}
# plot, ggplot
ggplot(regions) +
  geom_sf(aes(fill = pct_water)) +
  scale_fill_distiller(
    "pct_water", palette = "Spectral", direction=1,
    guide = guide_legend(title = "% water", reverse=T)) +
  theme_bw() +
  ggtitle("% Water by US Region")
```

## Regions: recalculate area

So far you've used the `ALAND` column for area of the state. But what if you were not provided the area and needed to calculate it? Because the `states` are in geographic coordinates, you'll need to either transform to an equal area projection and calculate area, or use geodesic calculations. Thankfully, the `sf` library provides area calculations with the `st_area()` and uses the `geosphere::distGeo()` to perform geodesic calculations (ie trigonometric calculation accounting for the spheroid nature of the earth). Since the `states` data has the unusual aspect of a z dimension, you'll need to first remove that with the `st_zm()` function.

```{r recalc-regions-area, warning=F, message=F}
library(geosphere)
library(units)

regions = states %>%
  mutate(
    water_m2 = AWATER %>% set_units(m^2),
    land_m2  = geometry %>% st_zm() %>% st_area()) %>%
  group_by(region) %>%
  summarize(
    water_m2 = sum(water_m2),
    land_m2  = sum(land_m2)) %>%
  mutate(
    pct_water = water_m2 / land_m2)

# table
regions %>%
  st_set_geometry(NULL) %>%
  arrange(desc(pct_water))
```

```{r plot-regions-area, warning=F, message=F}
# plot, ggplot
ggplot(regions) +
  geom_sf(aes(fill = as.numeric(pct_water))) +
  scale_fill_distiller(
    "pct_water", palette = "Spectral", direction=1,
    guide = guide_legend(title = "% water", reverse=T)) +
  theme_bw() +
  ggtitle("% Water by US Region")
```

## Challenge: project & recalculate area {.challenge}

Use `st_transform()` with a [USA Contiguous Albers Equal Area Conic Projection](http://spatialreference.org/ref/esri/usa-contiguous-albers-equal-area-conic/) that minimizes distoration, and then calculate area using the `st_area()` function.

### Answers {.solution} 

```{r recalc-regions-transform}
library(geosphere)
library(units)

# Proj4 of http://spatialreference.org/ref/esri/usa-contiguous-albers-equal-area-conic/
crs_usa = '+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs'

regions = states %>%
  st_transform(crs_usa) %>%
  mutate(
    water_m2 = AWATER %>% set_units(m^2),
    land_m2  = geometry %>% st_zm() %>% st_area()) %>%
  group_by(region) %>%
  summarize(
    water_m2 = sum(water_m2),
    land_m2  = sum(land_m2)) %>%
  mutate(
    pct_water = water_m2 / land_m2)

# table
regions %>%
  st_set_geometry(NULL) %>%
  arrange(desc(pct_water))

# plot, ggplot
ggplot(regions) +
  geom_sf(aes(fill = as.numeric(pct_water))) +
  scale_fill_distiller("pct_water", palette = "Spectral") +
  theme_bw() +
  ggtitle("% Water (geodesic) by US Region")
```

## Key Points {.keypoints}

- The `sf` package can take advantage of chaining spatial operations using the `%>%` operator.
- Data manipulation functions in `dplyr` such as `group_by()`, `summarize()` and `mutate()` work on `sf` objects.
- Area can be calculated a variety of ways. Geodesic is preferred if starting with geographic coordinates (vs projected).
