--- 
title: "NPS Workshop on Interactive Reporting with R"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [style/zotero_nps.bib, style/packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: ecoquants/nps-r-workshop
description: "These materials are for a 2-day workshop on Interactive Reporting with R for the National Park Service (NPS). The output format is using `bookdown::gitbook()`."
---

```{r include=FALSE, eval=F}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'style/packages.bib')
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(leaflet)
library(sf)
opts_chunk$set(echo = F)
```

Location: [401 W Hillcrest Dr, Thousand Oaks, CA 91360](https://goo.gl/maps/tJ7YHFwwoyq)<br>
Dates: November 13-14, 2018

# Overview

This is the landing site for our workshop materials. Google Drive (folders, docs, slides, sheets, etc.) allows us to be agile with adapting content on the fly. The rest of the content in this site is generated using the techniques taught: R + markdown = Rmarkdown, git, Github, etc.

## Google Drive Content

- [**workshop/**](https://drive.google.com/drive/u/1/folders/19mPz-w4AHHn0X3Jje5XvC3n5c1SnMlJC)
    - [**agenda**](https://docs.google.com/document/d/1mS_DcEoCZDN8-O4h7DeBdHzsk-5FkailM58Hm56jp6c/edit#)
    - [**notes**](https://docs.google.com/document/d/1bVvVOZISIT9ZyEsOGmqr0LccQaHl9gcl-EoraJBg7Gw/edit#)
    - [**presentations/**](https://drive.google.com/drive/u/1/folders/1MNolN6cPbbf1k33Qg1L4YLAvVdv7cBuR)
    - [**data/**](https://drive.google.com/drive/u/1/folders/1ge660eD5b7OinQSKSLjrCSKyFk3bJE2L)

## Resources

- [R for Data Science](http://r4ds.had.co.nz/)
- [Spatial Data Analysis and Modeling with R](http://rspatial.org)
- [Introduction to GIS - slides](https://docs.google.com/presentation/d/1P3ITbMxVsYcxbzUXLbDo8DLBs7d4nC_20eilIuaB0X0/present)

