--- 
title: "Reproducible Reporting with R (R³) for marine ecosystem indicators"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [style/zotero_nps.bib, style/packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: ecoquants/nps-r-workshop
description: "These materials are for a 1-day workshop on Reproducible Reporting with R for the Working Group on the Northwest Atlantic Regional Sea (WGNARS) at Dartmouth SMAST on March 16, 2020. The output format is using `bookdown::gitbook()`."
---

```{r include=FALSE, eval=F}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'style/packages.bib')
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(leaflet)
library(sf)
opts_chunk$set(echo = F)
```

# Overview {-}

These materials are for a 1-day workshop on Reproducible Reporting with R for the Working Group on the Northwest Atlantic Regional Sea (WGNARS) at Dartmouth SMAST on March 16, 2020.

- _**Location**_: [School for Marine Science & Technology (SMAST), UMass Dartmouth](https://www.umassd.edu/smast/)<br>
[706 S Rodney French Blvd, New Bedford, MA 02744](https://goo.gl/maps/LWMUfGmGsYdcAVo58)<br>
- _**Date**_: Monday, March 16, 2020

This is the landing site for our workshop materials. Google Drive (folders, docs, slides, sheets, etc.) allows us to be agile with adapting content on the fly. The rest of the content in this site is generated using the techniques taught: R + markdown = Rmarkdown, git, Github, etc.

## Google Drive Content {-}

- [**workshop/**](https://drive.google.com/drive/u/1/folders/19mPz-w4AHHn0X3Jje5XvC3n5c1SnMlJC)
    - [**agenda**](https://docs.google.com/document/d/1mS_DcEoCZDN8-O4h7DeBdHzsk-5FkailM58Hm56jp6c/edit#)
    - [**notes**](https://docs.google.com/document/d/1bVvVOZISIT9ZyEsOGmqr0LccQaHl9gcl-EoraJBg7Gw/edit#)
    - [**presentations/**](https://drive.google.com/drive/u/1/folders/1MNolN6cPbbf1k33Qg1L4YLAvVdv7cBuR)
    - [**data/**](https://drive.google.com/drive/u/1/folders/1ge660eD5b7OinQSKSLjrCSKyFk3bJE2L)

## Resources {-}

- [R for Data Science](http://r4ds.had.co.nz/)
- [Spatial Data Analysis and Modeling with R](http://rspatial.org)
- [Introduction to GIS - slides](https://docs.google.com/presentation/d/1P3ITbMxVsYcxbzUXLbDo8DLBs7d4nC_20eilIuaB0X0/present)

