---
pagetitle: Spatial
output:
  html_document:
    pandoc_args: [
      "--number-offset=3"]
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
shelf(
  htmltools, mitchelloharawild/icons)
```

# Map

## Learning Objectives {.unnumbered .objectives}

- TBD

## Read Spatial Data

Open your `r3-exercises.Rproj` to launch RStudio into that project and set the working directory.

Create a new Rmarkdown file (RStudio menu File > New file > Rmarkdown...) called `map.Rmd`. Insert headers like last time followed by Chunks of R code according to the examples provided below.

I'll be copy/pasting during the demonstration but I encourage you to type out the text to enhance understanding.

### Install packages

```{r}
# require() is like library() except returns FALSE if missing (vs error)
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
# librarian::shelf() is like library() except installs package if missing, 
#   even from Github if include owner/repo
shelf(
  NOAA-EDAB/ecodata,
  sf)
# librarian before version 1.8.1, got error:
#   download from 'https://api.github.com/repos/NOAA-EDAB/ecodata/tarball/HEAD' failed
```

## Get spatial data

```{r}
ecodata::epu_sf

epu_sf <- ecodata::epu_sf %>% 
  st_transform(4326)
```

So we see a geometry list column.
```{r}
class(epu_sf)
# "sf"         "data.frame"
g1 <- epu_sf$geometry[1]
# see in Environment pane, expand g1
```

```{r}
plot(epu_sf)
plot(epu_sf["EPU"])
```
Where in the world is this?

```{r}
shelf(mapview)

mapview(epu_sf)
```

```{r}
shelf(leaflet)

leaflet() %>% 
  #addTiles() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addPolygons(data = epu_sf)
```

## Group by

sf is "tidy"

## Extract from erddap

* [ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Monthly - Data Access Form](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html)


* [CoastWatch ERDDAP](https://coastwatch.pfeg.noaa.gov/erddap/index.html): search for "SST":
  * [`jplMURSST41mday`: ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Monthly - Data Access Form](https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html)

```{r}
shelf(
  here,
  rerddap)

sst_gd_rds <- here("data/sst_gd.rds")

epu_bb <- st_bbox(epu_sf)
epu_bb

sst_info <- info('jplMURSST41mday')
sst_info

if (!file.exists(sst_gd_rds)){
  sst_gd <- griddap(
    sst_info,
    fields    = "sst",
    time      = c("2020-06-16", "2021-06-16"),
    longitude = epu_bb[c("xmin", "xmax")],
    latitude  = epu_bb[c("ymin", "ymax")])
  saveRDS(sst_gd, file = sst_gd_rds)  
}
sst_gd <- readRDS(sst_gd_rds)

sst_gd
names(sst_gd)
```

```{r}
shelf(
  dplyr,
  ggplot2,
  mapdata)

# coastline
coast <- map_data(
  "worldHires", 
  xlim = epu_bb[c("xmin", "xmax")],
  ylim = epu_bb[c("ymin", "ymax")],
  lforce = "e")

sst_df_last <- sst_gd$data %>%
    filter(time == max(time))

# summary(sst_last)

ggplot(
  data = sst_df_last, 
  aes(x = lon, y = lat, fill = sst)) +
  geom_polygon(
    data = coast,
    aes(x = long, y = lat, group = group), fill = "grey80") +
  geom_tile() +
  scale_fill_gradientn(
    colors = rerddap::colors$temperature, na.value = NA) +
  theme_bw() + 
  ylab("Latitude") + 
  xlab("Longitude") +
  ggtitle("Latest SST")
```

```{r}
shelf(
  purrr,
  raster,
  sp,
  tidyr)
select <- dplyr::select

sst_tbl <- tibble(sst_gd$data) %>%
  mutate(
    # round b/c of uneven intervals
    #   unique(sst_gd$data$lon) %>% sort() %>% diff() %>% table()
    #     0.0099945068359375 0.0100021362304688
    lon  = round(lon, 2),
    lat  = round(lat, 2),
    date = as.Date(time, "%Y-%m-%dT00:00:00Z")) %>%
  select(-time) %>% 
  filter(!is.na(sst)) # 13M to 8.8M rows

sst_tbl_mo <- sst_tbl %>% 
  nest(data = c(lat, lon, sst)) %>% 
  mutate(
    raster = purrr::map(data, function(x) {
      #browser()
      sp::coordinates(x) <- ~ lon + lat
      sp::gridded(x)     <- T
      raster::raster(x)
    }))

sst_stk <- raster::stack(sst_tbl_mo$raster)
names(sst_stk) <- strftime(sst_tbl_mo$date, "sst_%Y.%m")
raster::crs(sst_stk) <- 4326
```


```{r}
shelf(stringr)

epu_sst_avg <- raster::extract(sst_stk, epu_sf, fun = mean, na.rm = T)
epu_sst_sd  <- raster::extract(sst_stk, epu_sf, fun = sd, na.rm = T)

epu_sst_tbl <- rbind(
  epu_sst_avg %>% 
    as_tibble() %>% 
    cbind(
      EPU  = epu_sf$EPU,
      stat = "mean") %>% 
    pivot_longer(-c(EPU, stat)),
  epu_sst_sd %>% 
    as_tibble() %>% 
    cbind(
      EPU = epu_sf$EPU,
      stat = "sd") %>% 
    pivot_longer(-c(EPU, stat))) %>% 
  mutate(
    EPU  = as.character(EPU),
    date = as.double(str_replace(name, "sst_", ""))) %>% 
  select(-name) %>% 
  pivot_wider(
    names_from  = EPU, 
    values_from = value)
```


```{r}
shelf(dygraphs)

epu_sst_tbl %>% 
  filter(stat == "mean") %>% 
  select(-stat) %>% 
  dygraph()
```


## Further Reading {.unnumbered}

- R packages
  * [`mapview`](https://r-spatial.github.io/mapview/#:~:text=mapview%20provides%20functions%20to%20very,the%20geometries%20and%20their%20attributes.)
  * [`leaflet`](https://rstudio.github.io/leaflet/)
  * [`rerddap`](https://docs.ropensci.org/rerddap/articles/rerddap.html)